\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2013-11-15 doc/paper.tex

\subsection{Lattice method}
If ready.
\subsection{multiple effective models at different scales}
\section{Implementing new models}
\section{Adapting the code at the C++ level}
\section{Using individual modules in your own code}
\appendix
\section{Examples}
\subsection{NMSSM}
Detailed steps to create NMSSM model file.
\begin{enumerate}
\item Create new directory \code{mkdir NMSSM}
\item Move into this new directory \code{cd NMSSM}
\item copy the MSSM model file in \code{cp /path/to/SARAH/Models/MSSM/MSSM.m  NMSSM.m}
\item Edit NMSSM.m as follows
\begin{enumerate}
\item  Change Model name, author and date
\item Add singlet superfield to SuperFields
\begin{lstlisting}
SuperFields[[8]] = {s, 1, sR,    0, 1,  1, RpM};
\end{lstlisting}
\item Modify superpotential
\begin{lstlisting}
SuperPotential = Yu u.q.Hu - Yd d.q.Hd - Ye e.l.Hd + \[Lambda] S.Hu.Hd +  \[Kappa]/3 s.s.s ;
\end{lstlisting}
\item Add new VEV
\begin{lstlisting} DEFINITION[EWSB][VEVs]= 
  { {SHd0, {vd, 1/Sqrt[2]}, {sigmad, \[ImaginaryI]/Sqrt[2]},{phid,1/Sqrt[2]}},
    {SHu0, {vu, 1/Sqrt[2]}, {sigmau, \[ImaginaryI]/Sqrt[2]},{phiu,1/Sqrt[2]}},
    {SsR,  {vs, 1/Sqrt[2]}, {sigmaS, \[ImaginaryI]/Sqrt[2]},{phiS,1/Sqrt[2]}}
};
\end{lstlisting}
\item Add new VEV
\begin{lstlisting} 
  DEFINITION[EWSB][VEVs]= 
  { {SHd0, {vd, 1/Sqrt[2]}, {sigmad, \[ImaginaryI]/Sqrt[2]},{phid,1/Sqrt[2]}},
    {SHu0, {vu, 1/Sqrt[2]}, {sigmau, \[ImaginaryI]/Sqrt[2]},{phiu,1/Sqrt[2]}},
    {SsR,  {vs, 1/Sqrt[2]}, {sigmaS, \[ImaginaryI]/Sqrt[2]},{phiS,1/Sqrt[2]}}
};
\end{lstlisting}
\item  Add the singlino to the list of gauge eigenstate dirac fermions
\begin{lstlisting}
  DEFINITION[GaugeES][DiracSpinors]={
  Bino ->{fB, conj[fB]},
  Wino -> {fWB, conj[fWB]},
  Glu -> {fG, conj[fG]},
  H0 -> {FHd0, conj[FHu0]},
  HC -> {FHdm, conj[FHup]},
  Fd1 -> {FdL, 0},
  Fd2 -> {0, FdR},
  Fu1 -> {FuL, 0},
  Fu2 -> {0, FuR},
  Fe1 -> {FeL, 0},
  Fe2 -> {0, FeR},
  Fv -> {FvL,0}
  FS -> {FsR,conj[FsR]}   
};
\end{lstlisting}
\item Edit mass eigenstates to give how the gauge eigenstates mix into mass eigenstates.  Here we add to the cp even higgs a real singlet, to the CP odd the imaginary part of the singlet and finally to the neutralinos we add the singlino.

@@

{{ Description -> "Singlet-VEV", 
			 LaTeX -> "v_s",
             Value -> None, 
             LesHouches -> {NMSSMRUN,5},
             OutputName-> vS         }},
\end{lstlisting}

\end{enumerate}   
\subsection{\ESSM}
Detailed steps to create \ESSM model file.

\section{Tests}
Comparisons against existing spectrum generators to test for deviations, e.g. the tests against MSSM and NMSSM (and E6SSM?).

>>>>

\subsection{Lattice method}
If ready.
\subsection{multiple effective models at different scales}
\section{Implementing new models}
\section{Adapting the code at the C++ level}
\section{Using individual modules in your own code}
\appendix
\section{Examples}
\subsection{NMSSM}
Detailed steps to create NMSSM model file.
\begin{enumerate}
\item Create new directory \code{mkdir myNMSSM}
\item Move into this new directory \code{cd myNMSSM}
\item copy the MSSM model file in \code{cp /path/to/SARAH/Models/MSSM/MSSM.m  myNMSSM.m}
\item Edit myNMSSM.m as follows
\begin{enumerate}
\item  Change Model name, author and date
\item Add singlet superfield to SuperFields
\begin{lstlisting}
SuperFields[[8]] = {s, 1, sR,    0, 1,  1, RpM};
\end{lstlisting}
\item Modify superpotential
\begin{lstlisting}
SuperPotential = Yu u.q.Hu - Yd d.q.Hd - Ye e.l.Hd + \[Lambda] s.Hu.Hd +  \[Kappa]/3 s.s.s ;
\end{lstlisting}
\item Add new VEV
\begin{lstlisting} DEFINITION[EWSB][VEVs]= 
  { {SHd0, {vd, 1/Sqrt[2]}, {sigmad, \[ImaginaryI]/Sqrt[2]},{phid,1/Sqrt[2]}},
    {SHu0, {vu, 1/Sqrt[2]}, {sigmau, \[ImaginaryI]/Sqrt[2]},{phiu,1/Sqrt[2]}},
    {SsR,  {vs, 1/Sqrt[2]}, {sigmaS, \[ImaginaryI]/Sqrt[2]},{phiS,1/Sqrt[2]}}
};
\end{lstlisting}
\item Add new VEV
\begin{lstlisting} 
  DEFINITION[EWSB][VEVs]= 
  { {SHd0, {vd, 1/Sqrt[2]}, {sigmad, \[ImaginaryI]/Sqrt[2]},{phid,1/Sqrt[2]}},
    {SHu0, {vu, 1/Sqrt[2]}, {sigmau, \[ImaginaryI]/Sqrt[2]},{phiu,1/Sqrt[2]}},
    {SsR,  {vs, 1/Sqrt[2]}, {sigmaS, \[ImaginaryI]/Sqrt[2]},{phiS,1/Sqrt[2]}}
};
\end{lstlisting}
\item  Add the singlino to the list of gauge eigenstate dirac fermions
\begin{lstlisting}
  DEFINITION[GaugeES][DiracSpinors]={
  Bino ->{fB, conj[fB]},
  Wino -> {fWB, conj[fWB]},
  Glu -> {fG, conj[fG]},
  H0 -> {FHd0, conj[FHu0]},
  HC -> {FHdm, conj[FHup]},
  Fd1 -> {FdL, 0},
  Fd2 -> {0, FdR},
  Fu1 -> {FuL, 0},
  Fu2 -> {0, FuR},
  Fe1 -> {FeL, 0},
  Fe2 -> {0, FeR},
  Fv -> {FvL,0}
  FS -> {FsR,conj[FsR]}   
};
\end{lstlisting}
\item Edit mass eigenstates to give how the gauge eigenstates mix into mass eigenstates.  Here we add to the cp even higgs a real singlet, to the CP odd the imaginary part of the singlet and finally to the neutralinos we add the singlino.

@@

{{ Description -> "Singlet-VEV", 
			 LaTeX -> "v_s",
             Value -> None, 
             LesHouches -> {NMSSMRUN,5},
             OutputName-> vS         }},
\end{lstlisting}
\item Navigate to Flexible SUSY directory  \code{cd /path/to/FlexibleSUSY/}
\item Create Z3NMSSM directory 
\begin{lstlisting}
./createmodel --models=Z3NMSSM:myNMSSM
\end{lstlisting}

\end{enumerate}   
\subsection{\ESSM}
Detailed steps to create \ESSM model file.
\begin{enumerate}
\item Create new directory \code{mkdir myE6SSM}
\item Move into this new directory \code{cd myE6SSM}
\item copy the MSSM model file in \code{cp /path/to/SARAH/Models/MSSM/MSSM.m  myE6SSM.m}
\item Edit myE6SSM.m as follows
\begin{enumerate}
\item  Change Model name, author and date
\item Modify MSSM superfield definitions to include new charges.
\begin{lstlisting}
SuperFields[[1]] = {q, 3, {uL,  dL},     1/6, 2, 3, Qq, RpM};  
SuperFields[[2]] = {l, 3, {vL,  eL},    -1/2, 2, 1, Ql, RpM};
SuperFields[[3]] = {Hd,1, {Hd0, Hdm},   -1/2, 2, 1, QHd, RpP};
SuperFields[[4]] = {Hu,1, {Hup, Hu0},    1/2, 2, 1, QHu, RpP};
SuperFields[[5]] = {d, 3, conj[dR],    1/3, 1, -3, Qd, RpM};
SuperFields[[6]] = {u, 3, conj[uR],   -2/3, 1, -3, Qu, RpM};
SuperFields[[7]] = {e, 3, conj[eR],      1, 1,  1, Qe, RpM};
\end{lstlisting}
Note that in the \ESSM model the charges are fixed to be those of the $U(1)_N$ as shown in e.g. \cite{King:2005jy}.  However here we leave the charges unassigned at the Mathematica level, and will set them as parameters in the C++ code.  The charges could also simply be fixed here to simplify the C++ code.

\item Add Singlet Higgs superfield to SuperFields
\begin{lstlisting}
SuperFields[[8]] = {s, 1, sR,     0, 1,  1, Qs, RpP};
\end{lstlisting}\end{lstlisting}
\item Add inert Higgs-like superfields 
\begin{lstlisting}
SuperFields[[9]] = {H1I, 2, {H1I0, H1Im},  -1/2, 2, 1, QHu, RpP};
SuperFields[[10]] = {H2I, 2, {H2Ip, H2I0},   1/2, 2, 1, QHd, RpP};
SuperFields[[11]] = {sI, 2, sR,    0, 1,  1, Qs, RpM};
\end{lstlisting}
\item Add exotics to SuperFields
\begin{lstlisting}
SuperFields[[12]] = {Dx, 3, conj[DxR],  1/3, 1, -3, QDx, RpP};
SuperFields[[13]] = {Dxbar, 3, conj[DxbarR],  -1/3, 1, 3, QDxbar, RpP};
\end{lstlisting}
\item Add additional survival $SU(2)$ multiplets to SuperFields
\begin{lstlisting}
SuperFields[[14]] = {Hp, 1, conj[HpR],    0.5, 2,  1, QHp, RpP};
SuperFields[[15]] = {Hpbar, 1, conj[HpbarR],    -0.5, 2,  1, QHpbar, RpP};
\end{lstlisting}
\item Modify superpotential
\begin{lstlisting}
SuperPotential = Yu u.q.Hu - Yd d.q.Hd - Ye e.l.Hd + \[Lambda] S.Hu.Hd +  La S.H2I.H1I + \[Kappa] S.Dx.Dxbar;
\end{lstlisting}
\item Add new VEV
\begin{lstlisting} DEFINITION[EWSB][VEVs]= 
  { {SHd0, {vd, 1/Sqrt[2]}, {sigmad, \[ImaginaryI]/Sqrt[2]},{phid,1/Sqrt[2]}},
    {SHu0, {vu, 1/Sqrt[2]}, {sigmau, \[ImaginaryI]/Sqrt[2]},{phiu,1/Sqrt[2]}},
    {SsR,  {vs, 1/Sqrt[2]}, {sigmaS, \[ImaginaryI]/Sqrt[2]},{phiS,1/Sqrt[2]}}
};
\end{lstlisting}

\end{enumerate}

\end{enumerate}   
\section{Tests}
Comparisons against existing spectrum generators to test for deviations, e.g. the tests against MSSM and NMSSM (and E6SSM?).
